---
title: "Exam 1"
author: "Westley Cook"
date: "2/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading the three packages I know I'll need to use at various points;
# tidyverse for its dplyr and ggplot2 packages, janitor to clean names after
# reading in data, and gt to make a table on question three

library(tidyverse)
library(janitor)
library(gt)

```

```{r read-paris-data, include=FALSE}

# Reading in paris data (after making the "raw-data" subdirectory and moving the
# .csv file there) and running clean_names() just to be safe

paris <- read_csv("raw-data/paris.csv") %>% 
  clean_names()

```

### Question 1

```{r question1.1, echo=FALSE}

# This r chunk assigns two objects to place in-line for mad lib 1.

# Arranging the paris data in descending order of emissions, slicing the first
# row and pulling the name column to get the country name.

in_country <- paris %>% 
  arrange(desc(emissions)) %>% 
  slice(1) %>% 
  pull(name)

# I realized after doing the above that the sentence "the highest emissions are
# from [country]" would be more meaningful with a bit more context (without
# context, one might misunderstand the sentence to mean something like "the
# highest *average* emissions) so I ran the same code but pulled the year column
# this time and added it to the end of the mad lib to clarify that the highest
# single-year emissions in the data set came from China in 2005.

in_year <- paris %>% 
  arrange(desc(emissions)) %>% 
  slice(1) %>% 
  pull(year)

```

1. The highest emissions are from the country **`r in_country`** in
**`r in_year`**.

```{r question1.2, echo=FALSE}

# This r chunk assigns an object to place in-line for mad lib 2.

# Filtering paris once to take out the ratifying countries, then again to take
# out the approving countries. The resulting data was small enough that I could
# look through the 9 rows to see that all of the countries left had indicated
# either acceptance or accession, and that the only NA values left were in the
# signature column for the two countries indicating accession. I figured that
# the signature probably wasn't required for countries indicating accession, and
# thus left those two countries in the total count.

number_countries <- paris %>% 
  filter(kind != "Ratification") %>% 
  filter(kind != "Approval") %>% 
  count() %>% 
  pull(n)

# I also added an explanatory note in italics which will show up on the finished
# .html file, saying what it means for a country in this dataset to have neither
# ratified nor approved the treaty.

```

2. The number of countries that have neither ratified nor approved of the treaty
is **`r number_countries`**.
*This number includes only countries who indicated acceptance or accession.*

```{r question1.3, echo=FALSE}

# This r chunk assigns an object to place in-line for mad lib 3.

# This one tripped me up for a little bit, because I'd originally used
# read.csv() instead of read_csv() to read in the paris data, and it made the
# ratification and signature columns factors instead of dates. I had a hard time
# converting them into the file type I wanted, but eventually figured out via
# google that read_csv() wouldn't create that problem for me. Once I got the
# columns to be the appropriate type, the rest was easy.

# Filtering for just the ratifying countries, then creating a new variable
# ("gap") to show the number of days between signature and ratification for each
# ratifying country (couldn't do this with these variables when they were
# factors, but can do it easily with dates; data type matters!). Arranging the
# data in descending "gap" order to find the highest value, then slicing the
# first row and pulling gap.

longest_days <- paris %>% 
  filter(kind == "Ratification") %>% 
  mutate(gap = ratification - signature) %>% 
  arrange(desc(gap)) %>% 
  slice(1) %>% 
  pull(gap)
  
```

3. The longest number of days between signature and ratification (for a country
that has successfully ratified) is **`r longest_days`** days.

```{r question1.4, echo=FALSE}

# This r chunk assigns an object to place in-line for mad lib 4.

# Filtering the paris data for countries whose name contains "stan" (because I
# didn't know how to limit the str_detect() function to just those countries
# whose names END with "stan"). But the result of that filter was just seven
# rows, so I looked at all of them and manually checked that they all ended in
# "stan" (rather than have a "stan" somewhere in their name). At that point it
# was easy to pipe the filter to a count() function to tell me the number of
# rows, and pull the number of rows.

number_stan <- paris %>% 
  filter(str_detect(name, "stan"),
         kind == "Ratification") %>% 
  count() %>% 
  pull(n)

```

4. The number of ratifying countries whose name ends with “stan” is 
**`r number_stan`**.

```{r read-co2-data, include=FALSE}

# This r chunk reads in the WB co2 emissions data.

# In order to figure out how many lines to skip to get the first rows of actual
# data (which had column names) at the top, I first read in the data without
# skipping any lines and used glimpse() and head() to see in what row the column
# names were located; I then added skip = 4 to the read_csv() function to cut
# out the unnecessary rows. I also used tail() to make sure there weren't rows
# at the bottom which I wouldn't need.

# I added the col_types = cols() argument later, in order to make the message
# "Parsed with..." go away in the output (this maybe wasn't necessary, since I'm
# using include=FALSE in the r chunk, but I thought it was good practice).

# Used clean_names() to fix the numeric column names.

emissions <- read_csv("raw-data/WB_co2_emisions.csv", 
                      skip = 4,
                      col_types = cols()
                      ) %>% 
  clean_names()

```

```{r question1.5, echo=FALSE}

# This r chunk assigns an object to place in-line for mad lib 5.

# Same process here as was used above on similar questions. The fact that each
# year was its own column made this particular question one step shorter than it
# would've been otherwise, because I just arranged by descending order of x1960
# instead of having to first filter for year == 1960 and then arranging by
# descending order of emissions. (But I get that tidy data is better to work
# with for most other questions!)

highest_1960 <- emissions %>% 
  arrange(desc(x1960)) %>% 
  slice(1) %>% 
  pull(country_name)

```

5. In 1960, the country with the highest emissions was **`r highest_1960`**.

```{r question1.6, echo=FALSE}

# This r chunk assigns an object to place in-line for mad lib 6.

# Same overall process here as in other mad lib chunks (pulling a value and
# assigning it to an object). The differences this time were that I used
# summarize() create a new variable (avg2000) taking the mean of all values in
# the x2000 column (using na.rm = TRUE to discount the NA values), and that I
# used the round() function to trim the mean to just two digits after the
# decimal point.

avg_in_2000 <- emissions %>% 
  summarize(avg2000 = mean(x2000, na.rm = TRUE)) %>% 
  pull(avg2000) %>% 
  round(digits = 2)

```
  
6. The average emissions in 2000 was **`r avg_in_2000`** metric tons per capita.

<br>

### Question 2

```{r question2, echo=FALSE}

# This r chunk tidies up the emissions dataframe, joins it with the paris
# dataframe, and then creates a plot of North American countries' CO2 emissions
# over time.

# First, tidying the emissions data by pivoting all the year columns into a new
# column ("year") and assigning their values to another new column,
# "emissions_per_capita". I use the names_prefix = "x" argument to uniformly
# drop the x prefix from the years as they go from being column names to being
# values in the "year" column. I use the type_convert() function to tell R to
# guess the column types for me. This is important for two reasons: 1, it gets
# rid of the "Parsed with..." message that would otherwise show up, and 2, it
# formats the columns as appropriate types (which, as I learned after the
# read.csv() vs read_csv() headache on question 1 mad lib 3, is always
# important).

emissions_tidy <- emissions %>% 
  pivot_longer(cols = c(x1960:x2019),
             names_to = "year",
             values_to = "emissions_per_captita",
             names_prefix = "x") %>% 
  type_convert(col_types = cols())
  
# Now, using a full_join() to attach the paris data to the tidy emissions data
# by year and country code. I chose full_join() because I knew I would need
# emissions data from all available years to make my plot, and that meant I had
# to choose either left_join() or full_join() (an inner_join() or right_join()
# would cut out data from years that weren't in the paris dataframe). I tried
# both full_join() and left_join() to see what would be different between the
# two, and chose full_join() because it preserved 8 rows of data that
# left_join() didn't - and since I didn't know what those 8 observations were, I
# thought it would probably be better to keep them in my dataframe just in case.

joined <- emissions_tidy %>% 
  full_join(paris,
             by = c("year", 
                    "country_code" = "code")
            )

# In order to use the date_of_effect variable in my plot as stipulated by the
# question prompt, I'm assigning an object called "date" with the date of effect
# of the Paris Treaty for the United States, Canada, and Mexico. I checked each
# country's date of effect by first filtering for those three countries, and
# then, noting that they were all the same, thought the easiest way to grab that
# date and assign it to an object would be what I've done below: filter for just
# that date, then slice the first row and pull the value for date_of_effect.

date <- joined %>%
  filter(date_of_effect == "2016-11-04") %>%
  slice(1) %>%
  pull(date_of_effect)

# The following code makes a plot of emissions per capita for the three North
# American countries from 1960-2014

joined %>% 
  
# Filtering for the three countries I want, and taking out observations were
# emissions_per_capita is NA
  
  filter(country_name == "United States" |
           country_name == "Canada" |
           country_name == "Mexico",
         emissions_per_captita != "NA") %>% 

# Plotting emissions by year, grouping and coloring by country; the group =
# country_name is important for when I add a line to the points
  
    ggplot(
      aes(x = year,
          y = emissions_per_captita,
          color = country_name,
          group = country_name)
      ) +
  
# Making it a scatterplot and connecting the dots by country  
  
  geom_point() +
  geom_line() +
  
# Adding a line to show when the Paris Accord took effect. Using the "date"
# object I created from the date_of_effect variable, formatting it to take just
# the year and casting it as a numeric to match the type of the "year" column on
# the x axis so it will show up in the appropriate place.
  
  geom_vline(xintercept = as.numeric(format(date, "%Y")),
             linetype = "dotted") +
  
# Adding a title and labels for the x and y axes. The title has extra spaces in
# it to center it over the plot; I probably could've accomplished the same thing
# better in a different way, but this seemed to work fine.
  
  labs(title = "   CO2 Emissions of North American Countries, 1960 - 2014",
       y = "Metric Tons Per Capita",
       x = "Year") +
  
# Adding a nice-looking name for the legend and changing the legend order to it
# matches the order of the lines (from highest emissions to lowest)
  
  scale_color_discrete(name = "Country",
                       breaks = c("United States",
                                  "Canada",
                                  "Mexico")
                       ) +
  
# Customizing the x axis in size and adding labels for every decade (the default
# labels and spacing wasn't quite as clear as I wanted to be)
  
  scale_x_continuous(breaks = seq(1960,
                                  2020,
                                  by = 10),
                     labels = c("1960",
                                "1970",
                                "1980",
                                "1990",
                                "2000",
                                "2010",
                                "2020")
                     ) +
  
# I love theme_classic(). It executes several lines of code to take away the
# awful-looking default background, grid, plot outline, etc., all with just a
# single line of code.
  
  theme_classic() +
  
# Adding a label for the vertical line showing when the Paris Accord took
# effect. Took a bit of trial and error to get the positioning to be exactly
# what I wanted it to be, but we got there.
  
  annotate("text",
           x = 2016, 
           y = 9.5, 
           label = "Year Paris Accord\nTook Effect (2016)",
           color = "gray28",
           size =3.5,
           angle = 90)

```

<br>

### Question 3

#### Potential Outcomes under the Rubin Causal Framework

The table below shows potential outcomes under the Rubin Causal Framework for
countries which ratified the Paris Accord. Suppose we had run an experiment in
which we assigned a treatment to all countries whose name begins with a vowel:
we gave each of them $1 billion to fight climate change.

(Assigning treatment to countries whose name begins with a
vowel may not be a very good random assignment mechanism and may even compromise
the results of our experiment, since there are far more countries whose names
begin with consonants and ideally we’d have ~50% of countries receiving
treatment and the remaining ~50% in the control group; nevertheless, suppose we
used it.)

The causal effect we are trying to measure is the difference in emissions for a
given country when it receives $1 billion compared to emissions for the same
country when it does *not* receive $1 billion. Measuring this causal effect is
difficult, because we only know one value per country: as you see in the table
(in which unknown values are marked by “?”) emissions for a “treated” country
are unknown in the “Control” column and vice versa. However, by making
assumptions to “fill in” the unknown values, we can approximate the causal
effect.

<br>

```{r question3, echo=FALSE}

# This r chunk first modifies the paris data to give "treatment" to all
# countries whose name begins with a vowel. It then makes a table out of the
# modified data, using one country for each letter of the alphabet.

# To give treatment, I first filter for only ratifying countries, then select
# the two variables I care about (name and emissions). I then copy the data from
# emissions into a new column, emissions_treatment, and rename the emissions
# column as emissions_control.

paris_experiment <- paris %>% 
  filter(kind == "Ratification") %>% 
  select(name, emissions) %>% 
  mutate(emissions_treatment = emissions) %>% 
  rename(emissions_control = emissions) %>% 
  
# Now for the fun part: I first use ifelse() to insert "?" into the
# emissions_treatment column for all countries whose name does NOT start with A,
# E, I, O, or U.
  
  mutate(emissions_treatment = ifelse(startsWith(name, "A") |
                                        startsWith(name, "E") |
                                        startsWith(name, "I") |
                                        startsWith(name, "O") |
                                        startsWith(name, "U"),
                                      emissions_treatment,
                                      "?")
         ) %>% 
  
# I then modify the function slightly to insert "?" into the emissions_control
# column for all countries whose name DOES start with a vowel.
  
  mutate(emissions_control = ifelse(startsWith(name, "A") |
                                        startsWith(name, "E") |
                                        startsWith(name, "I") |
                                        startsWith(name, "O") |
                                        startsWith(name, "U"),
                                     "?",
                                     emissions_control)
         )

# Now I take the modified paris data (paris_experiment) and make a table out of
# it

paris_experiment %>%
 
# I wanted to show one country for each letter of the alphabet, and the best way
# I knew to do that was to filter explicitly for the countries I wanted. I'm
# sure there's a more concise way to code this, but I don't know what it is...
# So I first made a table with all 177 countries, and then just took the first
# one from each letter and put them in this filter for the final table.
  
   filter(name == "Afghanistan" |
           name == "Bahamas" |
           name == "Cabo Verde" |
           name == "Democratic People's Repulbic of Korea" |
           name == "Ecuador" |
           name == "Finland" |
           name == "Gabon" |
           name == "Haiti" |
           name == "India" |
           name == "Jamaica" |
           name == "Kazakhstan" |
           name == "Lao People's Democratic Republic" |
           name == "Madagascar" |
           name == "Namibia" |
           name == "Oman" |
           name == "Pakistan" |
           name == "Qatar" |
           name == "Republic of Korea" |
           name == "Samoa" |
           name == "Tajikistan" |
           name == "Uganda" |
           name == "Vanuatu" |
           name == "Zambia") %>% 
  
# Making a nice-looking table, and adding a title and subtitle  
  
  gt() %>% 
  tab_header(title = "Potential Outcomes for Paris Experiment on Emissions",
             subtitle = "For Selected Countries") %>% 
  
# Adding spanner label for columns emissions_control and emissions_treatment  
  
  tab_spanner(label = "Potential Outcomes",
              columns = vars(emissions_control,
                             emissions_treatment)
              ) %>% 
  
# Adding one footnote to explain which countries are showing up in the table  
  
  tab_footnote(footnote = "Only one country is shown for each letter of the 
               alphabet (the first one in alphabetical order)",
               locations = cells_column_labels("name")
               ) %>% 
  
# Adding another footnote to explain within the table what the treatment is and
# to which countries it was given
  
  tab_footnote(footnote = "Treatment is receiving $1 billion to fight climate 
               change, and was given to all countries whose name begins with 
               a vowel",
               locations = cells_column_labels("emissions_treatment")
               ) %>% 
  
# Giving the columns pretty labels  
  
  cols_label(name = "Country",
             emissions_control = "Emissions Under Control",
             emissions_treatment = "Emissions Under Treatment") %>% 
  
# Center-aligning the values in the two emissions columns (which leaves the
# country names left-aligned)
  
  cols_align(align = "center",
             columns = vars(emissions_control,
                            emissions_treatment))

```